Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-03-31 WSSSPE1_report/WSSSPE1_summary_report.tex
Second paragraph in Implementing Policy.

The Workshop contributions in this group were concerned with the aspect of how implementation of best practices and other recommendations for improving scientific software sustainability could be promoted. Specifically, if scientific software is to become more sustainable, corresponding policies and guidelines need to be such that the scientific community can follow and implement them. This is considerably more challenging that it might seem at first, because in the reality of science today resources, both financial and personnel, that could be devoted to implementation are very limited, and the reward system does not encourage scientists to do so. Furthermore, implementing sustainability-targeting policies and guidelines often takes a variety of specialized software engineering expertises.

% To create sustainable software, we need to look at some key issues.
% First, we need to acknowledge that as a software package matures, 
% we will face new types of problems, and 
% we need to plan for these throughout a package’s life cycle. 
% Therefore, making scientific software sustainable means that we define policies and guidelines that the scientific community can follow and implement.
% But the reality of science today is that we have limited resources and rewards to encourage people to follow these policies and guidelines.
% Implementing policy often takes specialized expertise in software engineering.

% In light of these issues, I’ll be discussing several papers that were contributed to the workshop. 
% Some of these papers focus on specific facets of software design that are not often addressed in scientific software development, such as 
% API governance and 
% Software Security, and 
% some of the papers discuss strategies to implement all of the different facets of sustainable software design in the framework of scientific software.

>>>>

The workshop contributions in this group were concerned with the aspect of how implementation of best practices and other recommendations for improving scientific software sustainability could be promoted. Specifically, if scientific software is to become more sustainable, corresponding policies and guidelines need to be such that the scientific community can follow and implement them. This is considerably more challenging that it might seem at first, because in the reality of science today resources, both financial and personnel, that could be devoted to implementation are very limited, and the reward system does not encourage scientists to do so. Furthermore, implementing sustainability-targeting policies and guidelines often takes a variety of specialized software engineering expertises, which are not necessarily found in a single engineer, and much less so in a domain scientist cross-trained in programming. Adding to the policy implementation challenges, applicable sustainability-promoting practices and guidelines will change through a software project's lifecycle, in particular as it gains maturity.

Two of the papers in this group focus on specific facets of software design that are important factors in a project's sustainability but are often addressed only late in the scientific software development cycle, if at all: Wolski et al~\cite{} look at API governance, and Heiland et al~\cite{} discuss maturity models for software security. The other two papers discuss implementation strategies for science from the perspective of facilitating many or all facets of sustainanability-oriented software design for science:  Blanton and Lenhardt\cite{} contrast large projects that have software infrastructue development built-in with cross-training domain scientist PIs in software engineering best practices. Huang and Lapp~\cite{} discuss how various specialized software engineering skills could be turned into shared instrumentation with low barriers to access.
